Constant Error Tolerance Computational Fuzzy Extractors

### 2014-02-07 constantRate/ConstCompFuzzExt.tex
moved iris stuff to intro

\paragraph{Fuzzy Extractors}
Cryptography relies on secrets for that are read multiple times for key derivation and authentication. However, many sources with sufficient randomness  provide similar but not identical values at each reading~(prominent examples include biometrics and other human-generated data~\cite{daugman2004,zviran1993comparison,brostoff2000passfaces,ellison2000protecting,mayrhofer2009shake,monrose2002password},
physically unclonable functions~\cite{pappu2002physical,tuyls2006puf,gassend2002silicon,suh2007physical}, 
and quantum information~\cite{bennett1988privacy}). Turning similar readings into identical values is known as \emph{information reconciliation}; further converting those values into uniformly random secret strings is known as \emph{privacy amplification}~\cite{bennett1988privacy}.
Both of these problems have interactive and noninteractive versions.  In this paper, we are interested the noninteractive case, which can be used, in particular, by  a single user trying to produce the same key from multiple readings of a physical source at different times.  The primitive accomplishing information reconciliation and privacy amplification noninteractively is known as a \emph{fuzzy extractor}; it is defined information-theoretically in~\cite{DBLP:journals/siamcomp/DodisORS08}

Fuzzy extractors consist of a pair of algorithms: \gen takes a source value $w$, and produces a key $r$ and a public helper value $p$.  The second algorithm \rep takes this helper value $p$ and a close $w'$ to reproduce the original key $r$.  The security guarantee is that $r$ produces by \gen is close to uniform (information-theoretically \cite{DBLP:journals/siamcomp/DodisORS08} or computationally \cite{fuller2013computational}), even given $p$, as long as $w$ comes from a high-quality distribution (traditionally, any distribution with sufficient min-entropy $m$). The correctness guarantee is that $r$ will be correctly reproduced by \rep as long as $w'$ is no farther than $t$ from $w$ in some metric space (in this paper, we focus on the Hamming metric on strings over some alphabet $\mathcal{Z}$ of length $\ell$).

@@

However, some of the distributions with $\Huse\le 0$ are prime candidates use in authentication. \lnote{examples here}. No current fuzzy extractor construction is known to work for them.

@@

Both constructions are based on  obfuscation of point programs~\cite{canetti1997towards}.  A point program $I_w(x)$ outputs $1$ if $x=w$ and $0$ otherwise.  Intuitively, an obfuscated version of a program reveals nothing past its input and output behavior.  For a point program, this means hiding all partial information about the point $w$. 
We need a strong version of point obfuscation, which is secure under composition.  
However, we note that for even this strong version, obfuscation of point programs is much simpler and more efficient than general obfuscation. In particular, it is achievable under particular number-theoretic assumptions~\cite{bitansky2010strong}. It can be made very efficient under a strong assumption on cryptographic hash functions~\cite{canetti1998perfectly}.

@@

\paragraph{Limitations} In order to achieve security with $\Huse\le 0$, our alphabet must be super-polynomial in the security parameter.  Our constructions are applicable for sources where many bits are expected to err consecutively, followed by many consecutive correct bits~(known as burst errors~\cite{gilbert1960capacity}).  Constructing a computational fuzzy extractor with the above advantages and a small alphabet is an open problem.  Furthermore, we may leak sensitive information about the underlying source $W$.  In \consref{cons:first construction} it is possible to learn the value of individual blocks $W_i$.  Our two constructions are not noisy point obfuscation for any reasonable definition of obfuscation.  Dodis and Smith~\cite{DBLP:conf/stoc/DodisS05} achieve a notion of noisy point obfuscation when $\Huse>>0$ using a information-reconciliation component that hides all partial information.  Constructing noisy point obfuscation for all sources remains an open problem.

The second construction does not require a large alphabet for security.  However, because of its lower error tolerance, $\Huse \le 0$ only for large alphabets (and if $\Huse>0$, then there are good information-theoretic constructions known~\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08}).
It would be nice to get information-theoretic security, but current techniques seem to work for all distributions, which we know is impossible.

\lnote{old stuff below}

Fuzzy extractors have not yielded key derivation for all noisy sources.
There is tension between security and error-tolerance.  The key should be uniform to an observer but nearby $w'$ should map to the same key.  By nearby, we mean that $\dis(w, w')\leq t$ for some metric $\dis$.  In the information-theoretic realm, increasing $t$ means a decrease in the strength of the key $r$.  This is because a fuzzy extractor can be converted into an error correcting code~\cite[Appendix C]{DBLP:journals/siamcomp/DodisORS08}.  The length of the key $r$ is dictated by size of the best code that corrects $t$ errors.  Error-correcting codes are well studied and we have tight bounds on this quantity. % how many codewords may be in a code that corrects $t$ errors.

We consider the Hamming metric.  For the Hamming metric, there exists a distribution $W$ where the key derived from $W$ may be no longer than the difference between the starting entropy of the source and logarithm of the number of error patterns to be corrected~(and we have constructions of fuzzy extractors that nearly meet this bound). For the remainder of this paper we call this difference the minimum usable entropy, denoted $\Huse$.  This represents a practical problem as important biometrics have $\Huse \le 0$~(see \apref{sec:iris no key}).

 %let $|B_t|$ be the number of points within distance $t$.  The length of the key $|r| < \Hoo(W) - \log |B_t|$~(see~\cite[Appendix C]{DBLP:journals/siamcomp/DodisORS08} for the precise conditions).  For the remainder of this paper, we denote this quantity by $\Huse = \Hoo(W) - \log |B_t|$.

To achieve security when $\Huse \le 0$, some restriction on the source distribution $W$ is necessary~(minimal conditions are discussed in~\apref{sec:minimal conditions}).  However, the tools for constructing information-theoretic fuzzy extractors have not yielded natural restrictions.  Fuller, Meng, and Reyzin define a computationally-secure version of fuzzy extractors~\cite{fuller2013computational}.  Computational fuzzy extractors produce pseudorandom instead of truly random keys.  A computational fuzzy extractor may exist for a larger class of distributions with $\Huse\le 0$ than is possible information-theoretically.

In this work, we construct computational fuzzy extractors when $\Huse\le 0$ for large classes of distributions.
In the computational setting, a key may be expanded once it is long enough to serve as input to a computational extractor.  We focus on providing a reasonable length key when $\Huse\le 0$.

The remainder of this paper is organized as follows: we cover notation and background on obfuscation and error correcting codes in \secref{sec:preliminaries}, describe computational fuzzy extractors in \secref{sec:fuzzy extractors}, and present our two constructions in Sections \ref{sec:construction} and \ref{sec:sampling} respectively.

@@

\label{sec:iris no key}
As an example, we focus on the human iris.  The iris is believed to be the best biometric for high security applications~\cite{prabhakar2003biometric}.  The best estimate for iris entropy is $250$ bits~\cite{daugman2004}.  Daugman uses specialized wavelets to derive a $2048$ bit string called an iris code.  There has been considerable research since the work of Daugman~\cite{daugman2004}.  However, These works do not dramatically alter the entropy estimate or error rate necessary for a reasonable false reject rate.  We use the Daugman's parameters for our calculations.

The precise number of errors that must be tolerated depends on the desired false reject rate (how often the correct key is produced).  For a false reject rate of $\le 80\%$, a $t$ of approximately $205$ is required.%\footnote{Daugman  improves his true accept rate using a masking vector, where bits that are not considered ``trustworthy'' are excluded from the comparison.  The comparison between the two readings is then made using bits not excluded by either image.  It is not clear how to extend into the fuzzy extractor setting.  Even with the masking vector $\Huse$ is negative for irises.}  
We have the following calculation for $\Huse$:
\begin{align*}
\Huse &= \Hoo(W) - \log |B_t|\\
&= 249 - \log \sum_{i=0}^{205} {2048 \choose i} = -707.
\end{align*}
Thus, it is impossible to output a key using a fuzzy extractor that outputs the same strength key for all distributions with the same entropy and error tolerance~(see \apref{sec:minimal conditions} for minimal conditions for fuzzy extractor security).
Irises are believed to be the strongest biometric but applying current fuzzy extractors to irises yields no meaningful security.

>>>>

\paragraph{Fuzzy Extractors}
Cryptography relies on long-term secrets for key derivation and authentication. However, many sources with sufficient randomness to form long-term secrets provide similar but not identical values at repeated readings~(prominent examples include biometrics and other human-generated data~\cite{daugman2004,zviran1993comparison,brostoff2000passfaces,ellison2000protecting,mayrhofer2009shake,monrose2002password},
physically unclonable functions~\cite{pappu2002physical,tuyls2006puf,gassend2002silicon,suh2007physical}, 
and quantum information~\cite{bennett1988privacy}). Turning similar readings into identical values is known as \emph{information reconciliation}; further converting those values into uniformly random secret strings is known as \emph{privacy amplification}~\cite{bennett1988privacy}.
Both of these problems have interactive and noninteractive versions.  In this paper, we are interested in the noninteractive case, which is useful for a single user trying to produce the same key from multiple readings of a physical source at different times.  
 A \emph{fuzzy extractor} is the primitive that accomplishes both information reconciliation and privacy amplification noninteractively; fuzzy extractors are defined information-theoretically in~\cite{DBLP:journals/siamcomp/DodisORS08}.

Fuzzy extractors consist of a pair of algorithms: \gen takes a source value $w$, and produces a key $r$ and a public helper value $p$.  The second algorithm \rep takes this helper value $p$ and a close $w'$ to reproduce the original key $r$.  The security guarantee is that $r$ produces by \gen is close to uniform (information-theoretically \cite{DBLP:journals/siamcomp/DodisORS08} or computationally \cite{fuller2013computational}), even given $p$, as long as $w$ comes from a high-quality distribution (traditionally, any distribution with sufficient min-entropy $m$). The correctness guarantee is that $r$ will be correctly reproduced by \rep as long as $w'$ is no farther than $t$ from $w$ in some metric space (in this paper, we focus on the Hamming metric on length $\ell$ strings over some alphabet $\mathcal{Z}$).

@@

Prime candidate sources for authentication have $\Huse\le 0$.  As an example, the iris is believed to be the best biometric for high security applications~\cite{prabhakar2003biometric}.  The best estimate for iris entropy is $249$ bits~\cite{daugman2004}.  Daugman uses specialized wavelets to derive a $2048$ bit string called an iris code.  The precise number of errors that must be tolerated depends on the desired false reject rate (how often the correct key is produced).  For a false reject rate of $\le 80\%$, a $t$ of approximately $205$ is required.  We have the following calculation for $\Huse$:
\[
\Huse = \Hoo(W) - \log |B_t|
= 249 - \log \sum_{i=0}^{205} {2048 \choose i} = -414.
\]
Note there is considerable subsequent research~\cite{gentile2009slic,gentile2009efficient,rathgeb2011combining} to Daugman's work but it affects the quantity not the quality of the above bound.  No current fuzzy extractor is known to work for irises.\footnote{The work of Hao et al.~\cite[Section 4.3]{hao2006combining} provides a similar analysis but underestimates the number of possible error patterns; their calculation is $\Huse \approx 44$.}

@@

Both constructions are based on  obfuscation of point programs~\cite{canetti1997towards}.  A point program $I_w(x)$ outputs $1$ if $x=w$ and $0$ otherwise.  Intuitively, an obfuscated version of a program reveals nothing past its input and output behavior.  For a point program, this means hiding all partial information about the point $w$. 
We need a strong version of point obfuscation, which is secure under composition.  
However, we note that for even this strong version, obfuscation of point programs is much simpler and more efficient than general obfuscation \lnote{omer, could you put in some references?}. In particular, it is achievable under particular number-theoretic assumptions~\cite{bitansky2010strong}. It can be made very efficient under a strong assumption on cryptographic hash functions~\cite{canetti1998perfectly}.

@@

\paragraph{Open Problems}
Both constructions require a large alphabet $\mathcal{Z}$---one whose size is more than polynomial in the security parameter.\footnote{Alternatively, in case of burst errors~\cite{gilbert1960capacity}, multiple symbols from a small alphabet can be combined into a single symbol from a large alphabet.}  Although the sample-then-obfuscate construction remains secure with a small alphabet, because of its lower error tolerance, $\Huse \le 0$ only for large alphabets.  For small alphabets where $\Huse>0$,  there are good information-theoretic constructions known~\cite[Section 5]{DBLP:journals/siamcomp/DodisORS08}.
Constructing a computational fuzzy extractor with the above advantages and a small alphabet is an open problem.

Using information-theoretic fuzzy extractors with additional privacy properties, Dodis and Smith~\cite[Section 5]{DBLP:conf/stoc/DodisS05} construct program obfuscators for the program $I_w(x)$ that tests if $x$ is within distance $t$ of $w$. The obfuscation is secure as long as $w$ comes from a distribution of sufficient min-entropy; in particular, the entropy must be high enough so that $\Huse>0$. Our constructions do not provide obfuscators for proximity queries, because they leak more information than whether $x$ is within distance $t$ of $w$ (for example, they may provide some information about the actual distance or about which coordinates agree). Constructing an efficient obfuscator for proximity queries when $\Huse<0$ is an open problem.

\medskip
The remainder of this paper is organized as follows: we cover notation and background on obfuscation and error correcting codes in \secref{sec:preliminaries}, describe computational fuzzy extractors in \secref{sec:fuzzy extractors}, and present our two constructions in Sections \ref{sec:construction} and \ref{sec:sampling} respectively.

@@

\label{sec:iris no key}
As an example, we focus on the human iris.  The iris is believed to be the best biometric for high security applications~\cite{prabhakar2003biometric}.  The best estimate for iris entropy is $249$ bits~\cite{daugman2004}.  Daugman uses specialized wavelets to derive a $2048$ bit string called an iris code.  There has been considerable research since the work of Daugman~\cite{daugman2004}.  These works do not dramatically alter the entropy estimate or error rate necessary for a reasonable false reject rate.  We use the Daugman's parameters for our calculations.

The precise number of errors that must be tolerated depends on the desired false reject rate (how often the correct key is produced).  For a false reject rate of $\le 80\%$, a $t$ of approximately $205$ is required.%\footnote{Daugman  improves his true accept rate using a masking vector, where bits that are not considered ``trustworthy'' are excluded from the comparison.  The comparison between the two readings is then made using bits not excluded by either image.  It is not clear how to extend into the fuzzy extractor setting.  Even with the masking vector $\Huse$ is negative for irises.}  
We have the following calculation for $\Huse$:
\begin{align*}
\Huse &= \Hoo(W) - \log |B_t|\\
&= 249 - \log \sum_{i=0}^{205} {2048 \choose i} = -414.
\end{align*}
Thus, it is impossible to output a key using a fuzzy extractor that outputs the same strength key for all distributions with the same entropy and error tolerance~(see \apref{sec:minimal conditions} for minimal conditions for fuzzy extractor security).
Irises are believed to be the strongest biometric but applying current fuzzy extractors to irises yields no meaningful security.
